Canopy leaf area of an evergreen Eucalyptus woodland does not respond to elevated

### 2015-05-04 manuscript.Rmd
abstract edits

Leaf area, quantified by the leaf area index (*L*), is a crucial driver of forest productivity, water use, and energy balance. Because *L* responds to environmental drivers, it can represent an important feedback to climate change, but its responses to rising CO~2~ and water availability of unmanaged forests have been poorly quantified. We studied canopy leaf area dynamics for `r nummon` months in a native evergreen *Eucalyptus* woodland exposed to Free Air CO~2~ Enrichment (the EucFACE experiment), in a temperate (hot summer) climate where water limitation is common. We hypothesized that, because of expected higher productivity and increased water-use efficiency, *L* should increase with elevated [CO~2~]. We estimated *L* from diffuse canopy transmittance, and measured monthly leaf litter production. Contrary to expectation, *L* did not respond to elevated [CO~2~] in this timeframe. We found that *L* varied between `r rangeLAI[1]` and `r rangeLAI[2]`, but did not show a clear annual pattern. Instead, dynamics of *L* showed a quick increase after heavy rainfall, and a steady during periods of low rainfall. Leaf litter production was correlated to changes in *L*, both during periods of decreasing *L* (when no leaf growth occurred), and during periods of increasing *L* (active shedding of old foliage when new leaf growth occurred). Leaf lifespan, estimated from mean *L* and total annual litter production, was significantly increased under elevated [CO~2~] (p `r pval(LLtest$p.value)`). We discuss these results in the context of modelling phenology of evergreen woodlands in response to environmental drivers.

>>>>

Canopy leaf area, quantified by the leaf area index (*L*), is a crucial driver of forest productivity, water use, and energy balance. Because *L* responds to environmental drivers, it can represent an important feedback to climate change, but its responses to rising CO~2~ and water availability of unmanaged forests have been poorly quantified. We studied canopy leaf area dynamics for `r nummon` months in a native evergreen *Eucalyptus* woodland exposed to Free Air CO~2~ Enrichment (the EucFACE experiment), in a temperate (hot summer) climate where water limitation is common. We hypothesized that, because of expected stimulation of productivity and  water-use efficiency, *L* should increase with elevated [CO~2~]. We estimated *L* from diffuse canopy transmittance, and measured monthly leaf litter production. Contrary to expectation, *L* did not respond to elevated [CO~2~] in this timeframe. We found that *L* varied between `r rangeLAI[1]` and `r rangeLAI[2]` across the study period, but did not show a clear annual pattern. Instead, dynamics of *L* showed a quick increase after heavy rainfall, and a steady during periods of low rainfall. Leaf litter production was correlated to changes in *L*, both during periods of decreasing *L* (when no leaf growth occurred), and during periods of increasing *L* (active shedding of old foliage when new leaf growth occurred). Leaf lifespan, estimated from mean *L* and total annual litter production, was significantly longer under elevated [CO~2~] (p `r pval(LLtest$p.value)`). Our main finding that *L* was not responsive to elevated CO~2~ are consistent with other forest FACE studies, but contrast the positive response commonly predicted by ecosystem models.
